---
title: "R - Assignment 3"
format: 
  html: 
    toc: true
editor: visual
---

### Instruction

1.  *Download the Rmarkdown file at [this link](R_submission.Rmd)*

2.  *Open the submission file using Rstudio and write R codes below each question to answer the question. To insert a code chunk, use Ctrl + Alt + I.*

3.  *Once you are done answering all the question, Knit the file (Use: Ctrl + Shift + K or Click to Knit -\> Knit to pdf) to convert the Rmarkdown file into a pdf file.*

4.  *Submit the pdf/Word to Canvas.*

------------------------------------------------------------------------

Please refer to the class notes to do the following problems. 

[Class Slides](note3.html)

## Problem 1

Given the following three time series. All of these series are monthly series (Thus, `frequency = 12`)

[Series 1](data/series1.csv)

```{r, eval=FALSE, echo=FALSE}
library(forecast)
y1 = round(ts(c(1:100)+10*cos(c(1:100))+rnorm(100, sd = 1), frequency = 12),2)
write.csv(y1, 'series1.csv', row.names = FALSE)
plot(hw(y1))
```


[Series 2](data/series2.csv)

[Series 3](data/series3.csv)


```{r, eval=FALSE, echo=FALSE}

y = read.csv('data/series1.csv')
y = ts(y, frequency = 12)
plot(y)

```


Apply the appropriate exponential smoothing model for each series to make a forecast for the next value of the series. Explain your reason when choosing the model. Plot the forecast for each series. 

## Problem 2

Find a time series and apply an appropriate exponential smoothing model for the series to make a forecast for the next value of the series. Explain your reason when choosing the model. Plot the forecast. 


---