---
title: "R - Assignment 4"
format: 
  html: 
    toc: true
editor: visual
---

### Instruction

1.  *Download the Rmarkdown file at [this link](ayu_pod_submission.Rmd)*

2.  *Open the submission file using Rstudio and write R codes below each question to answer the question. To insert a code chunk, use Ctrl + Alt + I.*

3.  *Once you are done answering all the question, Knit the file (Use: Ctrl + Shift + K or Click to Knit -\> Knit to pdf) to convert the Rmarkdown file into a pdf file.*

4.  *Submit the pdf/Word to Canvas.*

------------------------------------------------------------------------

## Question 1

-   Use the following codes to plot the Airline Passenger data and its additive decomposition

```{r, eval=FALSE}
data("AirPassengers")
plot(AirPassengers)
ourDecomposition <- decompose(AirPassengers, type="additive")
plot(ourDecomposition)
```

-   Change `additive` to `multiplicative` to plot the multiplicative decomposition

-   Plot the [Electric_Production](data/Electric_Production.csv) and its decomposition (*Notice: You need to use function `ts` to create the time series first*)

-   Plot the [Electric_Production](data/Electric_Production.csv) from 1985 to 1995 and its decomposition

```{r, eval=FALSE, echo=FALSE}
d =  read.csv('data/Electric_Production.csv')
t2 = ts(d$IPG2211A2N, start = 1985, end = 1995, frequency = 12)
ourDecomposition <- decompose(t2, type="additive")
plot(ourDecomposition)
```

-   Include your codes in an Rmarkdown or Quarto, Knit it to either pdf or Word document. Submit the file to Canvas.


## Question 2

- Use the following codes to plot the Airline Passenger data and its STL decomposition

```{r, eval=FALSE}
data("AirPassengers")
plot(AirPassengers)
ourDecomposition <- stl(AirPassengers, s.window = "periodic")
plot(ourDecomposition)
```

- Plot the [Electric Production](data/Electric_Production.csv) and its stl decomposition (*Notice: You need to use function `ts` to create the time series first*)

- Plot the [Electric Production](data/Electric_Production.csv) from 1985 to 1995 and its stl decomposition

```{r, eval=FALSE, echo=FALSE}
d =  read.csv('data/Electric_Production.csv')
t2 = ts(d$IPG2211A2N, start = 1985, end = 1995, frequency = 12)
ourDecomposition <- decompose(t2, type="additive")
plot(ourDecomposition)
```

- Plot the [US Retail Employment](data/us_retail_employment.csv) and its STL decomposition. 

- Find your own time series that has a seasonal component and plot the classical decomposition (as in Assignment 3) and the STL decomposition. 

## Question 3

1. Run the codes in the [slides](https://bryantstats.github.io/math475/slides/5.html#1) to reproduce the results. The links to the datasets used in the slides are as follows.

  - [rainfall_london.csv](https://bryantstats.github.io/math475/slides/rainfall_london.csv)
  
  - [skirts.csv](https://bryantstats.github.io/math475/slides/skirts.csv)
  
  - [souvenir.csv](https://bryantstats.github.io/math475/slides/souvenir.csv)

2. Find three examples of the below time series and make forecast using Exponential Smoothing

- Time series with no trend and no seasonality
- Time series with trend and no seasonality
- Time series with trend and seasonality

---